---
title: '2023 30-day Map Challenge • Day 14 • Europe'
author: 'Ryan Peek'
execute:
  echo: false
  warning: false
  message: false
format: 
  html:
    self-contained: true
    code-overflow: wrap
    code-fold: true
    code-tools: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    theme: journal #litera #default
    highlight-style: arrow
editor_options: 
  chunk_output_type: console
---

## Europe: Degree of River Regulation Comparison

Comparing Ireland rivers with Sweden rivers to assess the degree of regulation (how much alteration and dam infrastructure exists).


## Ireland Data

```{r}
#| label: the-data-irl
#| message: false
#| warning: false
#| echo: true
#| eval: true

library(geodata)
library(tidyverse)
library(hrbrthemes)
library(glue)
library(ggtext)
library(sf)
library(rnaturalearth)
library(tigris)
options(tigris_use_cache = TRUE)

library(patchwork)
library(colorspace)
sf::sf_use_s2(FALSE)
library(showtext)
font_add_google("Irish Grover")
font_add_google("Roboto Slab")
showtext_auto()

# Get Data ----------------------------------------------------------------

# get IRL
irl <- ne_states(country = c("Ireland"), returnclass = "sf")
nir <- ne_states(country = "United Kingdom", returnclass = "sf") |> 
  filter(region == "Northern Ireland")
ireland <- bind_rows(irl, nir)
#plot(ireland$geometry, border="gray50")
#plot(st_union(ireland$geometry), border="forestgreen", lwd=2, add=TRUE)

# get worldclim data for precip and elev
d_wclim <- worldclim_country(country = "ireland", 
                       var = "prec",
                       path = tempdir())
d_elev <- elevation_30s(country = "ireland",
                       path = tempdir())

# plot
terra::plot(mean(d_wclim), col=viridis::mako(256), plg = list(title = "Precipitation"))
plot(ireland$geometry, add=TRUE, border=alpha("white", 0.3))

terra::plot(d_elev, col=viridis::viridis(256), plg = list(title = "Elev"))
plot(ireland$geometry, add=TRUE, border=alpha("white", 0.3))

# raster data of elev (takes a while)
#library(elevatr)
#d_elev2 <- get_elev_raster(locations = ireland, z = 10, 
#                     clip = "locations")
#terra::plot(rast(d_elev2), col=viridis::mako(256), plg = list(title = "Elevation (m)"))
#plot(ireland$geometry, add=TRUE, border=alpha("white", 0.3))

```

## Get Degree of Regulation (DOR) from Ireland

From: HydroSHEDS and Global Reservoir and Dams (GRanD) database v1.1.

```{r}
#| label: process-data-irl
#| message: false
#| warning: false
#| echo: true
#| eval: false

# GET RIVERS and IDS
rivs_eu <- st_read("/Users/rapeek/Documents/spatial_data/ATLAS/HydroRIVERS_v10_eu.gdb/HydroRIVERS_v10_eu.gdb", "HydroRIVERS_v10_eu")
# crop to Ireland
rivs_ire <- rivs_eu[ireland,]
#plot(rivs_ire$Shape, lwd = rivs_ire$ORD_STR, col=alpha("blue4", 0.5))
#plot(ireland$geometry, border=alpha("orange", 0.5), add=TRUE)
rm(rivs_eu)

# Read in based on SQL and get only first few rows
rivs_atlas_db <- "/Users/rapeek/Documents/spatial_data/ATLAS/RiverATLAS_Data_v10.gdb/RiverATLAS_v10.gdb"
st_layers("/Users/rapeek/Documents/spatial_data/ATLAS/RiverATLAS_Data_v10.gdb/RiverATLAS_v10.gdb")
# read only first few rows
rivs_atlas <- st_read(dsn=rivs_atlas_db, 
                      query="select * from \"RiverATLAS_v10\" limit 5")

# read matching options (takes a min)
rivs_atlas <- st_read(dsn=rivs_atlas_db, query=glue("select HYRIV_ID,dor_pc_pva from \"RiverATLAS_v10\" where HYRIV_ID in ({glue_collapse(rivs_ire$HYRIV_ID[1:100], sep=',')}) limit 10"))

# k try all!
rivs_atlas <- st_read(dsn=rivs_atlas_db, query=glue("select HYRIV_ID,dor_pc_pva from \"RiverATLAS_v10\" where HYRIV_ID > {min(rivs_ire$HYRIV_ID)} and HYRIV_ID < {max(rivs_ire$HYRIV_ID)}"))

# this takes forever!! save out?
#write_rds(rivs_atlas, file ="data_out/rivs_atlas_irl.rds")
rivs_atlas <- read_rds("data_out/rivs_atlas_irl.rds")

# join back to sf data
identical(rivs_atlas$HYRIV_ID, rivs_ire$HYRIV_ID)
rivs_atlas <- st_drop_geometry(rivs_atlas)

rivers_ire <- left_join(rivs_ire, rivs_atlas)
#summary(rivers_ire$dor_pc_pva)

# basic plot
#plot(st_union(ireland), border=alpha("gray40", 0.5), col=alpha("forestgreen", 0.2), lwd=2)
#plot(rivers_ire$Shape, lwd = rivers_ire$ORD_STR*.3, col=alpha("blue4", 0.4), add=TRUE)
#plot(rivers_ire$Shape[rivers_ire$dor_pc_pva>0], col="orange", lwd=3, add=TRUE)
```

## Plot Ireland

```{r}
#| label: plot-data-irl
#| message: false
#| warning: false
#| echo: true
#| eval: false


# ggplot
library(ggnewscale)
ireland_bord <- st_union(ireland)
(gg_ire <- ggplot() + 
  geom_sf(data=ireland_bord, color=alpha("seagreen", 0.5), fill=alpha("darkgreen", 0.2), lwd=0.7) +
  # try mapping to size
  geom_sf(data=rivers_ire, aes(color=as.factor(ORD_STRA), size=as.factor(ORD_STRA)), show.legend = FALSE)+
  scale_color_manual(name = "",
    values = rev(c('#08306b', '#08519c', '#2171b5', '#4292c6', '#6baed6', alpha('#6baed6',0.7), alpha('#6baed6',0.4)))) +
  scale_size_manual(values=c("1" = 0.1, "2" = 0.2, "3" = .45, "4" = .7, "5" = 1.2, "6"=2, "7"=3.1)) +
  new_scale_colour() +
  geom_sf(data=rivers_ire |> filter(dor_pc_pva>0), aes(color=dor_pc_pva), lwd=2, alpha=0.9) +
  scale_color_viridis_c("DOR", option="C") +
  ggthemes::theme_map() +
  labs(title="Degree of River Regulation in Ireland",
       subtitle = "#30DayMapChallenge • Day-14 • Europe",
       caption = "R. Peek • Data: HydroSheds RiverATLAS") +
  theme(plot.title = element_text(family ="Irish Grover", size = 80, hjust=0.5),
        plot.subtitle = element_text(family="Irish Grover", size=50, hjust=0.5),
        plot.caption = element_text(family="Roboto Slab", size=38),
        legend.title = element_text(family = "Roboto Condensed", face="bold", size=34),
        legend.text = element_text(family = "Roboto Condensed", face="bold", size=34),
        legend.position = c(0.94, 0.79)))

# save
ggsave(filename = "figs/day14-europe-irlrivs.png", width = 8, height = 10, dpi=300, bg="white")

```

## Sweden Data

```{r}
#| label: the-data-swe
#| message: false
#| warning: false
#| echo: true
#| eval: true

font_add_google("Nanum Pen Script")
font_add_google("Roboto Slab")
showtext_auto()

# Get Data ----------------------------------------------------------------

# get IRL
sweden <- ne_states(country = c("Sweden"), returnclass = "sf")

plot(sweden$geometry, border="gray50")
plot(st_union(sweden$geometry), border="gold", lwd=2, add=TRUE)

# get worldclim data for precip and elev
d_wclim <- worldclim_country(country = "sweden", 
                       var = "prec",
                       path = tempdir())
d_elev <- elevation_30s(country = "sweden",
                       path = tempdir())

# plot
terra::plot(mean(d_wclim), col=viridis::mako(256), plg = list(title = "Precipitation"))
plot(sweden$geometry, add=TRUE, border=alpha("white", 0.3))

terra::plot(d_elev, col=viridis::viridis(256), plg = list(title = "Elev"))
plot(sweden$geometry, add=TRUE, border=alpha("white", 0.3))

# raster data of elev (takes a while)
#library(elevatr)
#d_elev2 <- get_elev_raster(locations = sweden, z = 10, 
#                     clip = "locations")
#terra::plot(rast(d_elev2), col=viridis::mako(256), plg = list(title = "Elevation (m)"))
#plot(sweden$geometry, add=TRUE, border=alpha("white", 0.3))

```

## Get Degree of Regulation (DOR) from Sweden

From: HydroSHEDS and Global Reservoir and Dams (GRanD) database v1.1.

```{r}
#| label: process-data-swe
#| message: false
#| warning: false
#| echo: true
#| eval: false

# GET RIVERS and IDS
rivs_eu <- st_read("/Users/rapeek/Documents/spatial_data/ATLAS/HydroRIVERS_v10_eu.gdb/HydroRIVERS_v10_eu.gdb", "HydroRIVERS_v10_eu")
# crop to sweden
rivs_swe <- rivs_eu[sweden,]
#plot(rivs_swe$Shape, lwd = rivs_swe$ORD_STR, col=alpha("blue4", 0.5))
#plot(sweden$geometry, border=alpha("orange", 0.5), add=TRUE)
rm(rivs_eu)

# Read in based on SQL and get only first few rows
rivs_atlas_db <- "/Users/rapeek/Documents/spatial_data/ATLAS/RiverATLAS_Data_v10.gdb/RiverATLAS_v10.gdb"
st_layers("/Users/rapeek/Documents/spatial_data/ATLAS/RiverATLAS_Data_v10.gdb/RiverATLAS_v10.gdb")

#range(rivs_swe$HYRIV_ID)

# this works!
rivs_atlas <- st_read(dsn=rivs_atlas_db, query=glue("select HYRIV_ID,dor_pc_pva from \"RiverATLAS_v10\" where HYRIV_ID > {min(rivs_swe$HYRIV_ID)} and HYRIV_ID < {max(rivs_swe$HYRIV_ID)}"))

# join back to sf data
#identical(rivs_atlas$HYRIV_ID, rivs_swe$HYRIV_ID)
rivs_atlas <- st_drop_geometry(rivs_atlas)
rivers_swe <- left_join(rivs_swe, rivs_atlas)

# basic plot
plot(st_union(sweden), border=alpha("gray40", 0.5), col=alpha("forestgreen", 0.2), lwd=2)
plot(rivers_swe$Shape, lwd = rivers_swe$ORD_STR*.3, col=alpha("blue4", 0.4), add=TRUE)
plot(rivers_swe$Shape[rivers_swe$dor_pc_pva>0], col="orange", lwd=3, add=TRUE)
```

## Plot sweden

```{r}
#| label: plot-data-swe
#| message: false
#| warning: false
#| echo: true
#| eval: false


# ggplot
library(ggnewscale)
sweden_bord <- st_union(sweden)
(gg_swe <- ggplot() + 
  geom_sf(data=sweden_bord, color=alpha("gold", 0.5), fill=alpha("gold", 0.5), lwd=0.7) +
  # try mapping to size
  geom_sf(data=rivers_swe, aes(color=as.factor(ORD_STRA), size=as.factor(ORD_STRA)), show.legend = FALSE)+
  scale_color_manual(name = "",
    values = rev(c('#08306b', '#08519c', '#2171b5', '#4292c6', '#6baed6', alpha('#6baed6',0.7), alpha('#6baed6',0.4)))) +
  scale_size_manual(values=c("1" = 0.1, "2" = 0.2, "3" = .45, "4" = .7, "5" = 1.2, "6"=2, "7"=3.1)) +
  new_scale_colour() +
  geom_sf(data=rivers_swe |> filter(dor_pc_pva>0), aes(color=dor_pc_pva), lwd=2, alpha=0.9) +
  scale_color_viridis_c("DOR", option="C") +
  ggthemes::theme_map() +
  labs(title="Sweden")+
       #subtitle = "#30DayMapChallenge • Day-14 • Europe",
       #caption = "R. Peek • Data: HydroSheds RiverATLAS") +
  theme(plot.title = element_text(family ="Irish Grover", size = 80, hjust=0.5),
        plot.subtitle = element_text(family="Nanum Pen Script", size=50, hjust=0.5),
        plot.caption = element_text(family="Roboto Slab", size=38),
        legend.title = element_text(family = "Roboto Condensed", face="bold", size=34),
        legend.text = element_text(family = "Roboto Condensed", face="bold", size=34),
        legend.position = c(0.94, 0.79)))

# save
ggsave(filename = "figs/day14-europe-swerivs.png", width = 8, height = 10, dpi=300, bg="white")

```


## Patchwork

```{r}


library(patchwork)
gg_ire + gg_swe

ggsave(filename = "figs/day14-europe-rivs-combined.png", width = 11, height = 8.5, dpi=300, bg="white")
```




