---
title: '2023 30-day Map Challenge • Day 03 • Polly-gons Lines'
author: 'Ryan Peek'
execute:
  echo: false
  warning: false
  message: false
code-fold: true
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 2
    toc-location: left
    toc-title: Contents
    code-overflow: wrap
    code-fold: true
    code-tools: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    theme: journal #litera #default
    highlight-style: arrow
editor_options: 
  chunk_output_type: console
---

## Polly-Gons!    

I do a lot of research and work with amphibians in my day job, find them fascinating, and thought a fun exercise would be to see what folks across the globe call larval frogs (tadpoles or pollywogs). There turns out to be many words with many roots!

The etymology may be rooted in Middle English, where `tadpole` is from *tadde* “toad” and *pol* “head”. For  `pollywog`, it's possibly from the *pol* "head" and *wyglen* “to wiggle”.

 > porwigle, pollywiggle, pollywoggle, pollywog, potladle, pollywig, purwiggy
 
## Data

Expert sources were the internet, wikipedia, and dictionaries.

```{r}
#| label: the-data
#| message: false
#| warning: false
#| echo: true

library(tidyverse)
dat <- tibble(
  "term"=c("pollywog","tadpole", "pollywog", "tadpole", "dikkopje","kikkervisje","rumpetroll","holopadd","rompetrol", "rompetrol", "tadpole","halakarta","haletudse","grodyngel","nuijapää", "ultu","renacuajo","tzimbalu", "kaulquappe",
           "kijanka","girino","mormoloc",
           "iribaş","пуголовок","girino","têtard",
           "tornapa","penbwl","ot teny", "kēdǒu", "tāwhana","kiluwiluwi"), 
  "country_loc"=c("Canada", "United States", "United States", "United States", "Netherlands","Netherlands", "Norway", "Sweden", "Bulgaria", "Romania", "New Zealand", "Iceland","Denmark", "Sweden","Finland", "Ecuador","Bolivia", "Nicaragua", "Germany",
               "Poland","Portugal","Romania","Turkey","Ukraine","Italy","France","Ireland","United Kingdom","Lao PDR", "China", "New Zealand","Zimbabwe"),
  "definition"=c("larvae", "larvae", "larvae","larvae",
                 "little fathead",
                 "little frog fish","rump troll", 
                 "tail toad", "tail troll","tail troll",
                 "larvae","tail toad", "tail toad", 
                 "frog fry","clubhead","tadpole",
                 "tadpole", "tadpole", 
                 "thick head slimy blob",
                 "paddle head","round time", "head hard","tadpole", "tadpole", "button head", "tadpole","turnip","tadpole","a little boy", "tadpole","bounce","tadpole"))

```


```{r}
#| label: plot-image
#| message: false
#| echo: false
#| warning: false

library(rnaturalearth)
library(sf)

wrld <- ne_countries(scale = "medium", returnclass = "sf")
st_crs(wrld)
# use eqearth projection
wrld <- st_transform(wrld, crs = "+proj=eqearth")

# select countries 
wrld_polly <- wrld |> #filter(sovereignt %in% dat$country_loc) |> 
  inner_join(dat, by=c("name"="country_loc")) |> 
  dplyr::relocate(.before=name, c(term, definition))

# plot
ggplot() +
  geom_sf(data=wrld, fill=alpha("gray40", 0.3)) +
  geom_sf(data=wrld_polly, fill=alpha(col="orange", 0.7)) +
  ggrepel::geom_label_repel(data=wrld_polly,
                           aes(label=term, geometry=geometry), 
                           stat = "sf_coordinates", force_pull = 0.5, 
                           force = .1, min.segment.length = .5, size=2, max.overlaps = 30,
                           segment.size  = 0.2,
    segment.color = "grey50") +
  ggspatial::annotation_north_arrow() +
  labs(
    x = NULL, y = NULL,
    title = "Polly-gons: Words for Tadpole",
    subtitle = "#30DayMapChallenge • Day-03 • Polly-gons",
    caption = "+proj=eqearth • R 4.3.1 • Ryan Peek \n <https://github.com/ryanpeek/2023_30-day-map-challenge/>",
  ) +
  hrbrthemes::theme_ipsum_rc(grid="") +
  theme(axis.text=element_blank())

ggsave(filename = "figs/day03-pollygons.png", width = 11, height = 8.5, dpi=300, bg = "white")


```
