---
title: '2023 30-day Map Challenge • Day 10 • North America'
author: 'Ryan Peek'
execute:
  echo: false
  warning: false
  message: false
format: 
  html:
    self-contained: true
    code-overflow: wrap
    code-fold: true
    code-tools: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    theme: journal #litera #default
    highlight-style: arrow
editor_options: 
  chunk_output_type: console
---

## North America: Longest Roads

Mapping the roads of North America seemed an interesting thing to think about trying, but I was curious what the longest road/highway/interstate may actually be. So this was an attempt. I love the old maps that stack river lengths or tree heights vertically on a side and then show additional information on the opposite side, almost like a sideways histogram. 

## Data

```{r}
#| label: the-data
#| message: false
#| warning: false
#| echo: true
#| eval: true

library(tidyverse)
library(hrbrthemes)
library(glue)
library(ggtext)
library(sf)
library(tigris)
options(tigris_use_cache = TRUE)
library(patchwork)
library(colorspace)
sf::sf_use_s2(FALSE)
library(showtext)
#extrafont::fonts() # get fonts avail
font_add_google("Peralta")
showtext_auto()

# Get Data ----------------------------------------------------------------

# get states
library(tidycensus)
us <- tigris::states() |> 
  filter(!STUSPS %in% 
           c("HI","AS","PR","MP","VI", "GU", "AK"))
#us <- shift_geometry(us,preserve_area = TRUE, position = "below")
plot(us$geometry)

# get state cnty boundaries
ca <- tigris::states() |> filter(STUSPS=="CA")
cnty <- tigris::counties(state="CA")

# get roads in state! (# this takes a minute)
# and it's a big file ~ 260MB
if(!file.exists("data_raw/ca_roads.rds")){
  ca_rds <- tigris::roads("CA", cnty$COUNTYFP)
  saveRDS(ca_rds, file = "data_raw/ca_roads.rds")
} else({
  ca_rds <- read_rds("data_raw/ca_roads.rds")
})

# add length
ca_rds <- ca_rds |> 
  mutate(rd_length_m = st_length(geometry)) |> 
  mutate(rd_length_mi = drop_units(set_units(rd_length_m, miles)))

# ALL US primary roads
# https://catalog.data.gov/dataset/tiger-line-shapefile-2019-nation-u-s-primary-roads-national-shapefile
library(units)
us_rds <- st_read("~/Downloads/tl_2019_us_primaryroads/tl_2019_us_primaryroads.shp") |> 
  mutate(rd_length_m = st_length(geometry)) |> 
  mutate(rd_length_mi = drop_units(set_units(rd_length_m, miles))) |> 
  st_transform(st_crs(us))

us_rds <- us_rds[us,]

# PLOT?
gg_rd1 <- ggplot() + 
  geom_sf(data=us, fill=NA, color="gray50") +
  geom_sf(data=us_rds, aes(color=rd_length_mi), lwd=0.5, show.legend = FALSE) +
  scale_color_viridis_c("Length (mi)", option="F", trans = "log1p", breaks = c(10, 25, 50,100,150))+
  coord_sf(label_graticule = "", datum = NA) +
  hrbrthemes::theme_ft_rc(
    base_family = "Peralta",
    caption_family = "Roboto Slab", 
    plot_title_size = 42, subtitle_size = 32, caption_size = 18) +
  labs(title="Longest Road Segments", 
       subtitle = "TIGER 2019 U.S. Primary Roads",
       caption = "R. Peek • Data: Tiger 2019 Primary Roads \n <https://github.com/ryanpeek/2023_30-day-map-challenge/>") +
  theme(plot.subtitle = element_text(color="gray80"),
        plot.caption = element_text(color="gray80"),
        legend.text = element_text(size = 16, color="gray80"))
  
gg_rd1

ggsave(plot=gg_rd1, filename="figs/day10-namerica-roads.png", width = 7, height = 10, units = "in", dpi = 300)

# ca map
gg_rd2 <- ggplot() + 
  geom_sf(data=cnty, fill=NA, color="gray50") +
  geom_sf(data=ca_rds |> filter(rd_length_mi > 5), aes(color=rd_length_mi, linewidth=rd_length_mi)) +
  scale_linewidth_continuous(range=c(0.2,2))+
  scale_color_viridis_c("Length (mi)", 
                        option="F", 
                        trans = "log1p", 
                        breaks = c(5, 10, 25, 50,100,150))+
  coord_sf(label_graticule = "", datum = NA) +
  hrbrthemes::theme_ft_rc(
    base_family = "Peralta",
    caption_family = "Roboto", 
    plot_title_size = 42, subtitle_size = 32, caption_size = 18) +
  labs(subtitle="California")+ 
  #     subtitle = "TIGER Road Data",
  #     caption = "R. Peek • Data: {tigris} \n <https://github.com/ryanpeek/2023_30-day-map-challenge/>") +
  guides(linewidth="none")+
  theme(plot.subtitle = element_text(color="gray80"),
        plot.caption = element_text(color="gray80"),
        legend.text = element_text(size = 16, color="gray80"))
gg_rd2  

# this is big and takes a while to render
ggsave(plot=gg_rd2, filename="figs/day10-namerica-ca-roads.png", width = 7, height = 10, units = "in", dpi = 300)
  
```

## Arrange Plots

```{r}
#| label: combine-data
#| message: false
#| warning: false
#| echo: true
#| eval: false

library(patchwork)
# standard
gall <- gg_rd2 +plot_spacer()+ gg_rd1 +
  patchwork::plot_layout(widths = c(1.4,-0.3, 2),
              guides = "collect") + 
  theme(plot.background = 
          element_rect(fill = "#252a32", 
                       colour = "#252a32"))

gall

ggsave(gall, filename = glue("figs/day10_northamerica_combined.png"), width = 11, height = 8.5, dpi = 300)


```

